---
title: "wbc population"
output:
  pdf_document: default
  html_document: default
date: "2025-11-13"
---
---
title: "Untitled"
output:
  pdf_document: default
  html_document: default
date: "2025-11-10"
---
```{r, echo=FALSE}
 library(readxl)
library(dplyr)
library(tidyverse)
Blood_count <- read_excel("smears_analysis_r_updated14.xlsx")
```




```{r, echo=FALSE}
Blood_count <- Blood_count %>%
  dplyr::rename(
    `total_cell` = `[Cells] c/ml`
  )

```


```{r, echo=FALSE}
Blood_count <- Blood_count %>%
  dplyr::rename(
    `Lympho%` = `Lymoph%`
  )

```

```{r, echo=FALSE}
Blood_count <- Blood_count %>%
  dplyr::rename(
    `Shark_Id` = `Sharks Id`
  )
```


```{r}
Blood_count <- Blood_count %>% 
  filter(!is.na(Shark_Id))
```

```{r, echo=FALSE}
Blood_count <- Blood_count %>%
  dplyr::rename(
    total_count        = `total count`
  )
```


```{r, echo=FALSE}
library(dplyr) # Make sure dplyr is loaded

# Fix: Explicitly pass the data frame as the first argument to rename()
Blood_count <- dplyr::rename(Blood_count, RBC_WBC = `rbc:wbc`)
```

```{r, echo=FALSE, results='hide'}
library(tidyverse)
library(car)        # For Levene's Test
library(lme4)
library(lmerTest)   # For p-values in mixed models
library(MuMIn)      # For R²
library(DHARMa)     # For model diagnostics

# Make sure your ratio column is numeric and factors are correctly set
Blood_count <- Blood_count %>%
  mutate(
    RBC_WBC = as.numeric(RBC_WBC),     # rename if your column is `rbc:wbc`
    week_factor = factor(week),
    Groupe_factor = factor(Groupe),
    Sex_factor = factor(Sex),
    Shark_Id_factor = factor(Shark_Id) # replace with your shark ID column name
  )

# ----- 📊 1. Levene’s Test (homogeneity of variance) -----
leveneTest(RBC_WBC ~ Groupe_factor * week_factor, data = Blood_count)

# ----- 🧪 2. Shapiro–Wilk Test (normality) -----
# Test normality of the raw RBC:WBC values
shapiro.test(Blood_count$RBC_WBC)
```


# shapiro and levene good



```{r, echo=FALSE, results='hide'}
library(lme4)
library(lmerTest) 
library(DHARMa)

# Note: Ensure the RBC_WBC column exists in Blood_count
Blood_count$Groupe[Blood_count$Groupe == "Control"] <- "Static"
# --- 1. LMM for RBC_WBC Ratio ---
lmm_RBC_WBC_ratio <- lmer(RBC_WBC ~ Groupe_factor * week_factor + Sex_factor +  (1 | Shark_Id_factor),
                         data = Blood_count)

summary(lmm_RBC_WBC_ratio)

# --- 2. Check residuals with DHARMa ---
sim_RBC_WBC <- simulateResiduals(lmm_RBC_WBC_ratio, n = 250)

print("\n--- DHARMa Residual Plots and Tests for RBC_WBC Ratio ---")
plot(sim_RBC_WBC)
anova(lmm_RBC_WBC_ratio)
# DHARMa Tests
testDispersion(sim_RBC_WBC)       # Over-/under-dispersion
testQuantiles(sim_RBC_WBC)        # Check quantile deviations
library(emmeans)
library(openxlsx)

# --- Post-hoc comparisons ---
emms_RBC_WBC <- emmeans(lmm_RBC_WBC_ratio, ~ Groupe_factor * week_factor)
pairs_RBC_WBC <- pairs(emms_RBC_WBC, adjust = "tukey") |> summary()

# --- Convert to data frame ---
pairs_RBC_WBC_df <- as.data.frame(pairs_RBC_WBC)
pairs_RBC_WBC_df <- pairs_RBC_WBC_df[, c("contrast", "estimate", "SE", "df", "t.ratio", "p.value")]

# --- Export to Excel ---
write.xlsx(pairs_RBC_WBC_df, file = "RBC_WBC_posthoc_pairs.xlsx", sheetName = "RBC_WBC_pairs", rowNames = FALSE)

# --- Residual diagnostics already done with DHARMa ---
# sim_RBC_WBC from your code can be plotted and tested
plot(sim_RBC_WBC)
testDispersion(sim_RBC_WBC)
testQuantiles(sim_RBC_WBC)
emms_week_main_rbcwbc <- emmeans(lmm_RBC_WBC_ratio, ~ week_factor)

# --- 2. Perform Pairwise Comparisons on the Marginal Means ---
# This tests the overall difference between the weeks (4 vs 6, 4 vs 8, 6 vs 8).
pairs_week_main_rbcwbc <- pairs(emms_week_main_rbcwbc, adjust = "tukey")

# --- 3. Display the Results ---

# Summarize the pairwise comparisons
pairs_summary_rbcwbc <- summary(pairs_week_main_rbcwbc)

print("--- Overall Pairwise Comparisons for Week Factor (RBC/WBC Ratio) ---")
print(pairs_summary_rbcwbc)
```



```{r}
library(dplyr)

Blood_count %>%
  group_by(Groupe_factor) %>%
  summarise(
    mean_ratio = mean(RBC_WBC, na.rm = TRUE),
    sd_ratio = sd(RBC_WBC, na.rm = TRUE),
    n = n()
  )
```





## GLR


```{r, echo=FALSE, results='hide'}
library(lme4)
library(lmerTest)
library(dplyr)
library(car) 


Blood_count <- Blood_count %>%
  mutate(
    GLR = `[Granulocyte]` / `[Lympho]`
  )

Blood_count <- Blood_count %>%
  mutate(GroupWeek = interaction(Groupe_factor, week_factor))

leveneTest(GLR ~ GroupWeek, data = Blood_count, center = median)
shapiro.test(Blood_count$GLR)

```




```{r, echo=FALSE}
# custom_colors <- c("Control" = "#009E73", "Winter" = "#56B4E9", "Summer" = "#D55")

# --- 0. Setup and Model Definition (As provided by user) ---
Blood_count$Groupe[Blood_count$Groupe == "Control"] <- "Static"

Blood_count <- Blood_count %>%
  mutate(
    GLR = `[Granulocyte]` / `[Lympho]`,
    # Ensure factors are defined if they weren't in the provided block
    Groupe_factor = factor(Groupe),
    week_factor = factor(week),
    Sex_factor = factor(Sex),
    Shark_Id_factor = factor(Shark_Id)
  )



# LMM fitted to GLR
lmm_GLR <- lmer(GLR ~ Groupe_factor * week_factor + Sex_factor + (1 | Shark_Id_factor),
                data = Blood_count)

print("--- Summary of GLR LMM ---")
summary(lmm_GLR)

# --- 1. Diagnostics (As provided by user) ---
library(DHARMa)

print("\n--- DHARMa Residual Diagnostics for GLR LMM ---")
res_total1 <- simulateResiduals(fittedModel = lmm_GLR, n = 1000)

plot(res_total1)

testUniformity(res_total1)
testDispersion(res_total1)
testZeroInflation(res_total1)
library(emmeans)
library(writexl)
library(dplyr)



print("--- GLR Post-hoc: Group vs. Group Comparisons (Within Weeks) ---")


emms_glr_group_by_week <- emmeans(
  lmm_GLR, 
  specs = ~ Groupe_factor | week_factor
)

# 2. Perform pairwise comparisons using Tukey adjustment (suitable for comparing groups)
pairs_glr_group_by_week <- pairs(
  emms_glr_group_by_week, 
  adjust = "tukey"
) |> summary()
anova(lmm_GLR)
# --- Convert to data frame ---
pairs_glr_df <- as.data.frame(pairs_glr_group_by_week)

print(pairs_glr_df)

# --- 3. Export to Excel ---
write_xlsx(pairs_glr_df, "GLR_Group_by_Group_Comparisons.xlsx")

print("\nSuccessfully exported: GLR_Group_by_Group_Comparisons.xlsx")
```


GLR includig GT =  lmer(GLR ~ Groupe_factor * week_factor + Sex_factor + (1 | Shark_Id_factor)
                                  Estimate Std. Error       df t value Pr(>|t|)    
(Intercept)                       1.80054    0.10057 46.39577  17.904  < 2e-16 ***
Groupe_factorSummer               0.05188    0.15094 50.47320   0.344  0.73248    
Groupe_factorWinter              -0.11634    0.13222 50.52627  -0.880  0.38308    
week_factor6                     -0.18080    0.12559 34.29794  -1.440  0.15905    
week_factor8                     -0.36536    0.12559 34.29794  -2.909  0.00632 ** 
Sex_factorMale                    0.02300    0.07409 16.39906   0.310  0.76018    
Groupe_factorSummer:week_factor6  0.37054    0.20251 34.29794   1.830  0.07600 .  
Groupe_factorWinter:week_factor6  0.31819    0.17761 34.29794   1.791  0.08205 .  
Groupe_factorSummer:week_factor8  0.26531    0.21102 35.78229   1.257  0.21680    
Groupe_factorWinter:week_factor8  0.36261    0.17761 34.29794   2.042  0.04895 *  
```{r, echo=FALSE}
# Plot
p_glr_emmeans <- ggplot(emms_df, aes(x = week_factor, y = emmean, group = Groupe_factor, color = Groupe_factor)) +
  geom_line(position = position_dodge(width = 0.5), linewidth = 1) +
  geom_point(position = position_dodge(width = 0.5), size = 3) +
  geom_errorbar(aes(ymin = emmean - SE, ymax = emmean + SE),
                width = 0.15, position = position_dodge(width = 0.5), linewidth = 0.8) +
  scale_color_manual(values = custom_colors) +
  labs(
    title = " GLR ratio with GT",
    x = "Week",
    y = "GLR (Granulocyte / Lymphocyte)",
    color = "Group"
  ) +
  theme_minimal(base_size = 14) +
  theme(legend.position = "bottom")

print(p_glr_emmeans)

library(openxlsx)  # for writing Excel files

# --- Post-hoc comparisons for GLR ---
pairs_GLR <- pairs(emms_GLR, adjust = "tukey") |> summary()

# --- Convert to data frame ---
pairs_GLR_df <- as.data.frame(pairs_GLR)

# --- Optional: reorder/select columns ---
pairs_GLR_df <- pairs_GLR_df[, c("contrast", "estimate", "SE", "df", "t.ratio", "p.value")]

# --- Export to Excel ---
write.xlsx(pairs_GLR_df, file = "GLR_posthoc_pairs.xlsx", sheetName = "GLR_pairs", rowNames = FALSE)

```


```{r echo=FALSE}
library(lme4)
library(lmerTest)
library(emmeans)
library(openxlsx)
library(DHARMa)
library(dplyr)


# --- 1. Prepare data ---
Blood_count <- Blood_count %>%
  mutate(
    GLR1 = `[granulocyte_no_gt]` / `[Lympho]`,
    Groupe_factor = factor(Groupe),
    week_factor = factor(week),
    Sex_factor = factor(Sex),
    Shark_Id_factor = factor(Shark_Id)
  )

# --- 2. Fit LMM ---
lmm_glr1 <- lmer(GLR1 ~ Groupe_factor * week_factor + Sex_factor + (1 | Shark_Id_factor),
                 data = Blood_count)

summary(lmm_glr1)

# --- 3. Residual diagnostics ---
res_glr1 <- simulateResiduals(lmm_glr1, n = 1000)
plot(res_glr1)
testUniformity(res_glr1)
testDispersion(res_glr1)
testZeroInflation(res_glr1)

# --- 4. Marginal means for each group * week ---
emms_glr1 <- emmeans(lmm_glr1, ~ Groupe_factor | week_factor)

# --- 5. Post-hoc pairwise comparisons within each week ---

library(openxlsx)  # for writing Excel files

# --- Post-hoc comparisons for GLR ---
# Calculate predicted marginal means for GLR1
emms_glr1 <- emmeans(lmm_glr1, ~ Groupe_factor * week_factor)

# Post-hoc pairwise comparisons (all pairs across groups and weeks)
pairs_glr1 <- pairs(emms_glr1, adjust = "tukey") |> summary()

# Convert to data frame
pairs_glr1_df <- as.data.frame(pairs_glr1)

# Optional: reorder columns
pairs_glr1_df <- pairs_glr1_df[, c("contrast", "estimate", "SE", "df", "t.ratio", "p.value")]

# Export to Excel
write.xlsx(pairs_glr1_df, file = "GLR1_posthoc_pairs_GROUPS.xlsx", sheetName = "GLR1_pairs", rowNames = FALSE)

```



```{r, echo=FALSE}
library(lme4)
library(lmerTest)
library(dplyr)
library(emmeans)
library(ggplot2)

# --- 3. Set custom colors for groups ---
custom_colors <- c("Control" = "#009E73", "Winter" = "#56B4E9", "Summer" = "#D55")

# --- 4. Compute predicted marginal means ---
emms_GLR1 <- emmeans(lmm_glr1, ~ Groupe_factor * week_factor)
emms_df1 <- as.data.frame(emms_GLR1)

# --- 5. Plot GLR over weeks for each group ---
p_glr_emmeans1 <- ggplot(emms_df1, aes(x = week_factor, y = emmean, group = Groupe_factor, color = Groupe_factor)) +
  geom_line(position = position_dodge(width = 0.5), linewidth = 1) +
  geom_point(position = position_dodge(width = 0.5), size = 3) +
  geom_errorbar(aes(ymin = emmean - SE, ymax = emmean + SE),
                width = 0.15, position = position_dodge(width = 0.5), linewidth = 0.8) +
  scale_color_manual(values = custom_colors) +
  labs(
    title = "Granulocyte / Lymphocyte ratio without GT",
    x = "Week",
    y = "GLR",
    color = "Group"
  ) +
  theme_minimal(base_size = 14) +
  theme(legend.position = "bottom")

print(p_glr_emmeans1)
```
```{r}

library(ggplot2)
library(dplyr)
library(tidyr)

# --- 0. Data Preparation and Factor Definition FIX ---

# 1. CRITICAL FIX 1: Change the name in the base character column
Blood_count$Groupe[Blood_count$Groupe == "Contrôle"] <- "Static"

Blood_count <- Blood_count %>%
  mutate(
    # Calculate Ratios (Assuming these column names are correct in your data)
    GLR = `[Granulocyte]` / `[Lympho]`,
    GLR1 = `[granulocyte_no_gt]` / `[Lympho]`,
    
    # Define factors for Sex and ID
    Sex_factor = factor(Sex),
    Shark_Id_factor = factor(Shark_Id),
    
    # 2. CRITICAL FIX 2: Define the plotting factor, enforcing Winter, Static, Summer order
    Groupe_factor = factor(
      Groupe, 
      levels = c("Winter", "Static", "Summer") 
    ),
    
    # 3. Create the descriptive factor for weeks
    week_label = factor(
      paste("Week", week), 
      levels = c("Week 4", "Week 6", "Week 8")
    )
  )

# Define the custom color palette (must be defined outside the pipeline)
custom_colors <- c(
  "Static" = "#D6C9AE", 
  "Winter"  = "#6F9FCC",
  "Summer"  = "#D97A66" 
)


# --- 1. Prep Data for Long Format and Filter GLR only ---
glr_only <- Blood_count %>%
  # Select only the GLR variable for plotting
  select(Groupe_factor, week_label, GLR) %>% 
  pivot_longer(
    cols = c(GLR),
    names_to = "Ratio_Type",
    values_to = "Value"
  )

# --- 2. Create the Final Box Plot for GLR ---

p_glr_only_final <- ggplot(
  glr_only, 
  aes(x = Groupe_factor, y = Value, fill = Groupe_factor)
) +
  # Create the box plot layer, showing outliers
  geom_boxplot(
    outlier.shape = 19, 
    outlier.colour = "black", 
    outlier.size = 2 
  ) +
  
  # --- Facet by Week, with descriptive label, moved to the BOTTOM ---
  facet_wrap(
    ~ week_label, 
    scales = "fixed", 
    strip.position = "bottom" 
  ) +
  
  # Apply the custom colors
  scale_fill_manual(values = custom_colors) +
  
  # Labels and Theme Adjustments (Clean Aesthetics)
  labs(
    title = "", # Removed Plot Title
    x = "", # Removed X-axis label
    y = "GLR Value (Including GT)",
    fill = "" # Removed Legend Title
  ) +
  
  theme_minimal(base_size = 14) +
  theme(
    # Remove X-axis text and ticks 
    axis.text.x = element_blank(),
    axis.ticks.x = element_blank(),
    
    # Adjust facet strip to be at the bottom and clean up background
    strip.placement = "outside",
    strip.background = element_blank(), 
    strip.text = element_text(face = "bold", size = 12), 
    
    legend.position = "bottom",
    panel.spacing = unit(0.5, "lines")
  )

print(p_glr_only_final)
```

```{r}
library(ggplot2)
library(dplyr)
library(tidyr)

# --- 0. Data Preparation and Factor Definition FIX ---

# CRITICAL FIX 1: Change the name in the base character column
Blood_count$Groupe[Blood_count$Groupe == "Contrôle"] <- "Static"

Blood_count <- Blood_count %>%
  mutate(
    # Calculate Ratios (Assuming these column names are correct in your data)
    GLR = `[Granulocyte]` / `[Lympho]`,
    GLR1 = `[granulocyte_no_gt]` / `[Lympho]`,
    
    # Define factors using the corrected 'Groupe' name
    Sex_factor = factor(Sex),
    Shark_Id_factor = factor(Shark_Id),
    
    # 2. Set the final desired plotting factor order: Winter, Static, Summer
    Groupe_factor = factor(
      Groupe, 
      levels = c("Winter", "Static", "Summer") 
    ),
    
    # 3. Create the descriptive factor for weeks
    week_label = factor(
      paste("Week", week), 
      levels = c("Week 4", "Week 6", "Week 8")
    )
  )

# Define the Antibody Color Palette (used for consistency)
custom_colors <- c(
  "Static" = "#D6C9AE", 
  "Winter"  = "#6F9FCC",
  "Summer"  = "#D97A66" 
)


# --- 1. Prep Data for Long Format and Filter GLR1 ---
glr1_only <- Blood_count %>%
  select(Groupe_factor, week_label, GLR1) %>% 
  pivot_longer(
    cols = c(GLR1),
    names_to = "Ratio_Type",
    values_to = "Value"
  )

# --- 2. Create the Final Box Plot for GLR1 ---

p_glr1_final_dot <- ggplot(
  glr1_only, 
  aes(x = Groupe_factor, y = Value, fill = Groupe_factor)
) +
  # Create the box plot layer, showing outliers as solid black dots
  geom_boxplot(
    outlier.shape = 19, 
    outlier.colour = "black", 
    outlier.size = 2 
  ) +
  
  # --- Facet by Week, with descriptive label, moved to the BOTTOM ---
  facet_wrap(
    ~ week_label, 
    scales = "fixed", 
    strip.position = "bottom" 
  ) +
  
  # Apply the custom colors
  scale_fill_manual(values = custom_colors) +
  
  # Labels and Theme Adjustments
  labs(
    title = "", # Removed Plot Title
    x = "", # Removed X-axis label
    y = "GLR1 Value (No GT)",
    fill = "" # Removed Legend Title
  ) +
  
  theme_minimal(base_size = 14) +
  theme(
    # Remove X-axis text and ticks 
    axis.text.x = element_blank(),
    axis.ticks.x = element_blank(),
    
    # Adjust facet strip to be at the bottom and clean up background
    strip.placement = "outside",
    strip.background = element_blank(), 
    strip.text = element_text(face = "bold", size = 12), 
    
    legend.position = "bottom",
    panel.spacing = unit(0.5, "lines")
  )

print(p_glr1_final_dot)
```


Glr signifcant lower for summer and control, signidcantd ecrease week 6 and 8. glr tend to increqse at week 8 for winter. no difference amoung sex